---
title: "Check bait/prey balance"
output: html_notebook
params:
full_pos_file: NULL
full_neg_file: NULL
train_pos_file: NULL
train_neg_file: NULL
train_maxflow_pos_file: NULL
train_maxflow_neg_file: NULL
validate_pos_file: NULL
validate_neg_file: NULL
validate_maxflow_pos_file: NULL
validate_maxflow_neg_file: NULL
test_pos_file: NULL
test_neg_file: NULL
test_maxflow_pos_file: NULL
test_maxflow_neg_file: NULL
method: NULL
partition_name: NULL
dataset: NULL
pos_limit: NULL
neg_limit: NULL
---

```{r info_table, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- paste("| **Parameter**        | Value           |",
              "|---------------|:-------------:|",
              paste0("| **Dataset** | ", params$dataset, " |"),
              paste0("| **Partition** | ", params$partititon_name, " |"),
              paste0("| **Method** | ", params$method, " |"),
              paste0("| **Positive Limit** | ", params$pos_limit, " |"),
              paste0("| **Negative Limit** | ", params$neg_limit, " |"),
              sep="\n")

cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```
```


```{r functions, echo=F}
library(ggplot2)
library(tidyverse)
library(networkD3)
library(dplyr)
library(R.utils)

get_scaling_value <- function(neg_file) {
  first_line <- readLines(neg_file, n = 1)
  N <- as.numeric(sub(".*Scaled:\\s*(\\d+).*", "\\1", first_line))
  return(N)
}

get_degree <- function(df) {
  bait_count <- as.data.frame(table(df["bait"]))
  colnames(bait_count) <- c("gene_name", "degree")
  bait_count["group"] <- "bait"

  prey_count <- as.data.frame(table(df["prey"]))
  colnames(prey_count) <- c("gene_name", "degree")
  prey_count["group"] <- "prey"

  rbind(bait_count, prey_count) %>% return
}

get_degree_diff <- function(df_degree_pos, df_degree_neg) {
  df_degree_diff <- merge(
    df_degree_pos,
    df_degree_neg,
    by = c("gene_name", "group"),
    how = "outer",
    suffixes = c("_pos", "_neg")
  )

  df_degree_diff[is.na(df_degree_diff)] <- 0
  return(df_degree_diff)
}

get_wasserstein <- function(df_delta, aimed_scaling) {
  max_degrees <- as.numeric(max(c(max(df_delta$degree_pos * aimed_scaling), max(df_delta$degree_neg))))
  pos_hist <- rep(0, max_degrees)
  neg_hist <- rep(0, max_degrees)
  pos_hist_values <- table(df_delta$degree_pos * aimed_scaling)
  pos_hist[as.numeric(names(pos_hist_values))] <- pos_hist_values
  neg_hist_values <- table(df_delta$degree_neg)
  neg_hist[as.numeric(names(neg_hist_values))] <- neg_hist_values
  pos_hist <- pos_hist / sum(pos_hist)
  neg_hist <- neg_hist / sum(neg_hist)

  cdf_pos <- cumsum(pos_hist)
  cdf_neg <- cumsum(neg_hist)
  wasserstein_distance <- sum(abs(cdf_pos - cdf_neg))


  return(list(wasserstein_distance, cdf_pos, cdf_neg))
}

plot_wasserstain <- function(df_delta, aimed_scaling) {
  prey_wasser_list <- df_delta %>%
    filter(group == "prey") %>%
    get_wasserstein(aimed_scaling)

  plot_df_prey <- data.frame(
    degree = seq_along(prey_wasser_list[[2]]),
    cum_degree_pos = prey_wasser_list[[2]],
    cum_degree_neg = prey_wasser_list[[3]],
    group = "prey"
  )

  bait_wasser_list <- df_delta %>%
    filter(group == "bait") %>%
    get_wasserstein(aimed_scaling)

  plot_df_bait <- data.frame(
    degree = seq_along(bait_wasser_list[[2]]),
    cum_degree_pos = bait_wasser_list[[2]],
    cum_degree_neg = bait_wasser_list[[3]],
    group = "bait"
  )

  stats_df <- data.frame(
    x = c(length(bait_wasser_list[[2]]) * .4, length(prey_wasser_list[[2]]) * .4),
    y = 0.1,
    group = c("bait", "prey"),
    label = c(
      sprintf("Wasserstein: %.3f", bait_wasser_list[[1]]),
      sprintf("Wasserstein: %.3f", prey_wasser_list[[1]])
    )
  )

  plot_df <- rbind(plot_df_prey, plot_df_bait)

  g <- ggplot(plot_df,
              aes(
                x = degree
              )) +
    ylim(0, 1) +
    geom_line(aes(y = cum_degree_pos, color = "positive cumsum")) +
    geom_line(aes(y = cum_degree_neg, color = "negative cumsum")) +
    facet_grid(~group, scales = "free") +
    geom_text(data = stats_df, aes(x = x, y = y, label = label),
              inherit.aes = FALSE) +
    theme_bw() +
    labs(
      x = "Positive degree",
      y = "Negative degree"
    )

  return(g)
}



plot_degree <- function(df_delta, scaling_factor) {
  df_delta %>%
    filter(group == "prey") %>%
  { cor(.$degree_neg, .$degree_pos, method = "spearman") } -> spear_prey

  df_delta %>%
    filter(group == "prey") %>%
  { cor(.$degree_neg, .$degree_pos, method = "pearson") } -> pear_prey

  df_delta %>%
    filter(group == "prey") %>%
  { lm(degree_pos ~ degree_neg, data = .)$coefficients[2] } -> prey_slope

  df_delta %>%
    filter(group == "bait") %>%
  { cor(.$degree_neg, .$degree_pos, method = "spearman") } -> spear_bait

  df_delta %>%
    filter(group == "bait") %>%
  { cor(.$degree_neg, .$degree_pos, method = "pearson") } -> pear_bait

  df_delta %>%
    filter(group == "bait") %>%
  { lm(degree_pos ~ degree_neg, data = .)$coefficients[2] } -> bait_slope

  bait_y_max <- df_delta %>%
    filter(group == "bait") %>%
  { max(.$degree_neg) }
  bait_x_max <- df_delta %>%
    filter(group == "bait") %>%
  { max(.$degree_pos) }


  stats_df <- data.frame(
    group = c("prey", "bait"),
    x = bait_x_max * .3,
    y = bait_y_max * 0.9,  # Bottom-left (0-1 scale)
    label = c(
      sprintf("Spear: %.3f\nPear: %.3f\nSlope^-1: %.3f\nScaling: %.0f:1",
              spear_prey, pear_prey, 1 / prey_slope, scaling_factor),
      sprintf("Spear: %.3f\nPear: %.3f\n1/Slope^-1: %.3f\nScaling: %.0f:1",
              spear_bait, pear_bait, 1 / bait_slope, scaling_factor)
    )
  )
  g <- ggplot(df_delta,
              aes(
                x = degree_pos,
                y = degree_neg,
                color = group
              )) +
    geom_point() +
    facet_grid(~group) +
    geom_abline(slope = scaling_factor, linetype = "dashed") +
    geom_text(data = stats_df, aes(x = x, y = y, label = label),
              inherit.aes = FALSE) +
    theme_bw() +
    labs(
      x = "Positive degree",
      y = "Negative degree"
    )
  return(g)
}
```
```{r input, echo=F}
full_pos <- read.table(params$full_pos_file, sep = "\t", header = TRUE)
full_neg <- read.table(params$full_neg_file, sep = "\t", header = TRUE)

train_pos <- read.table(params$train_pos_file, sep = "\t", header = TRUE)
train_neg <- read.table(params$train_neg_file, sep = "\t", header = TRUE)
train_maxflow_pos <- read.table(params$train_maxflow_pos_file, sep = "\t", col.names = c("bait", "prey"))
train_maxflow_neg <- read.table(params$train_maxflow_neg_file, sep = "\t", col.names = c("bait", "prey"), skip = 1)
aimed_scaleing_factor_train <- get_scaling_value(params$train_maxflow_neg_file)
df_degree_pos_train <- get_degree(train_maxflow_pos)
df_degree_neg_train <- get_degree(train_maxflow_neg)
df_delta_train <- get_degree_diff(df_degree_pos_train, df_degree_neg_train)

validate_pos <- read.table(params$validate_pos_file, sep = "\t", header = TRUE)
validate_neg <- read.table(params$validate_neg_file, sep = "\t", header = TRUE)
validate_maxflow_pos <- read.table(params$validate_maxflow_pos_file, sep = "\t", col.names = c("bait", "prey"))
validate_maxflow_neg <- read.table(params$validate_maxflow_neg_file, sep = "\t", col.names = c("bait", "prey"), skip = 1)
aimed_scaleing_factor_validate <- get_scaling_value(params$validate_maxflow_neg_file)
df_degree_pos_validate <- get_degree(validate_maxflow_pos)
df_degree_neg_validate <- get_degree(validate_maxflow_neg)
df_delta_validate <- get_degree_diff(df_degree_pos_validate, df_degree_neg_validate)

test_pos <- read.table(params$test_pos_file, sep = "\t", header = TRUE)
test_neg <- read.table(params$test_neg_file, sep = "\t", header = TRUE)
test_maxflow_pos <- read.table(params$test_maxflow_pos_file, sep = "\t", col.names = c("bait", "prey"))
test_maxflow_neg <- read.table(params$test_maxflow_neg_file, sep = "\t", col.names = c("bait", "prey"), skip = 1)
aimed_scaleing_factor_test <- get_scaling_value(params$test_maxflow_neg_file)
df_degree_pos_test <- get_degree(test_maxflow_pos)
df_degree_neg_test <- get_degree(test_maxflow_neg)
df_delta_test <- get_degree_diff(df_degree_pos_test, df_degree_neg_test)
```



## PPI proportions
```{r plot_ppi_proportions, echo=FALSE}
get_n_pairs <- function(file, drop = 0) {
  n_lines <- countLines(file)[[1]]
  return(n_lines - drop)
}


all_pos_npair <- nrow(full_pos)
train_pos_npair <- nrow(train_pos)
validation_pos_npair <- nrow(validate_pos)
test_pos_npair <- nrow(test_pos)
mf_train_pos_npair <- nrow(train_maxflow_pos)
mf_validation_pos_npair <- nrow(validate_maxflow_pos)
mf_test_pos_npair <- nrow(test_maxflow_pos)

positive_data <- data.frame(
  source = c("All", "All", "All", "Train", "Validation", "Test", "Train", "Validation", "Test", "All"),
  target = c("Train", "Validation", "Test", "MF_Train", "MF_Validation", "MF_Test", "Discarded", "Discarded", "Discarded", "Discarded"),
  value = c(
    train_pos_npair,
    validation_pos_npair,
    test_pos_npair,
    mf_train_pos_npair,
    mf_validation_pos_npair,
    mf_test_pos_npair,
    (train_pos_npair - mf_train_pos_npair),
    (validation_pos_npair - mf_validation_pos_npair),
    (test_pos_npair - mf_test_pos_npair),
    (all_pos_npair -
      train_pos_npair -
      validation_pos_npair -
      test_pos_npair)
  )
)
nodes <- data.frame(
  name = c(as.character(positive_data$source),
           as.character(positive_data$target)) %>% unique()
)

positive_data$IDsource <- match(positive_data$source, nodes$name) - 1
positive_data$IDtarget <- match(positive_data$target, nodes$name) - 1

# Make the Network
p <- sankeyNetwork(
  Links = positive_data, Nodes = nodes,
  Source = "IDsource", Target = "IDtarget",
  Value = "value", NodeID = "name",
  sinksRight = FALSE,
  width = 1200,
  height = 800,
  fontSize = 32, nodeWidth = 64)
p

```



## Train data
```{r plot_train}
suppressWarnings(plot_degree(df_delta_train, aimed_scaleing_factor_train))
plot_wasserstain(df_delta_train, aimed_scaleing_factor_train)
```

## Validation data
```{r val_plot}
suppressWarnings(plot_degree(df_delta_validate, aimed_scaleing_factor_validate))
plot_wasserstain(df_delta_validate, aimed_scaleing_factor_validate)
```


## Test data
```{r, test_plot}
suppressWarnings(plot_degree(df_delta_test, aimed_scaleing_factor_test))
plot_wasserstain(df_delta_test, aimed_scaleing_factor_test)
```

